Meta-analysis of fungicide application timing experiments for powdery mildew

### 2020-05-22 03_Prep_data.Rmd

The following code attempts to convert 'least square differences' (LSD) to sampling variance as per the method in [Nugugi et.al (2011)](https://apsjournals.apsnet.org/doi/10.1094/PHYTO-08-10-0221).
Ideas for using a T-critical value of 1.697 came from reading a [statisics-how-to website](https://www.statisticshowto.datasciencecentral.com/how-to-calculate-the-least-significant-difference-lsd/).
A [table of T-critical values](https://www.statisticshowto.datasciencecentral.com/tables/t-distribution-table/) was consulted where 30 degrees of freedom within experiment was used to find the 'T-crit' value of 1.697.

@@

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates) # degrees of freedom within groups

   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$yield_error[1]

   V_yield <-
      (PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[1] * (LSD /
                                                                            Tcrit) ^ 2) / 2 # method - Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates))
      } # reset Sv vector if this is the first loop

@@

      Sv[j] = V_yield /
         PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[j]

@@

   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$Y_error_type <-
      "SamplingVariance"

@@

for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$yield_error)) {
   if (j == 1) {
      Sv <-
         vector(length = length(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$yield_error))
   }# create empty vector at first iteration with length of j

   Sv[j] = (PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$yield_error[j]) ^
      2
}

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref, ]$Y_error_type <-
   "SamplingVariance"
```

@@

for (i in unique(PM_MB_means$trial_ref)) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i,]

@@

   for (j in unique(dat1$row_spacing)) {
      dat2 <- dat1[dat1$row_spacing == j,]
      controlY <-
         mean(dat2[dat2$fungicide_ai == "control", "grain_yield.t.ha."])

@@

      dat1[dat1$row_spacing == j,] <- dat2
   }
   PM_MB_means[PM_MB_means$trial_ref == i,] <- dat1
}

@@

dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303",]

@@

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)
PM_MB_means[PM_MB_means$trial_ref == "AM1303",] <- dat1

@@

```{r simpler_clusters_REMOVE}

slimmer_PM_dat <- slim_PM_dat[slim_PM_dat$spray_management != "Late" & 
               slim_PM_dat$spray_management != "Early_plus",]

@@

slimmer_PM_dat %>%
   ggplot(aes(x= grain_yield.t.ha, y = vi))+
   geom_point()
```

@@

```{r Var_MeanSquarePlot}
slimmer_PM_dat %>%
   ggplot(aes(x= grain_yield.t.ha, y = Y_Msquare))+
   geom_point()

@@

plot(grain_yield.t.ha. ~ AUDPC_m,data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))

@@

```{r AUDPCxTrial}
sm2lg <- PM_MB_means%>%
   group_by(trial_ref)%>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE))%>%
   arrange(TY)

for(i in sm2lg$trial_ref){
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if(any(is.na(dat1$grain_yield.t.ha.))){next}
   if(any(is.na(dat1$AUDPC_m))){next}

plot(grain_yield.t.ha. ~ AUDPC_m, 
     data = dat1, main = paste(dat1$location[1],(dat1$year[1])),
     ylim = c(0,max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)), 
     xlim = c(0,max(PM_MB_means$AUDPC_m, na.rm = TRUE)))
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
points(grain_yield.t.ha. ~ AUDPC_m, data = dat1[dat1$fungicide_ai == "control",], col = "red")

@@

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m,data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

@@

sm2lg <- PM_MB_means%>%
   group_by(trial_ref)%>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE))%>%
   arrange(TY)

for(i in sm2lg$trial_ref){
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if(any(is.na(dat1$grain_yield.t.ha.))){next}
   if(any(is.na(dat1$AUDPS_m))){next}

plot(grain_yield.t.ha. ~ AUDPS_m, 
     data = dat1, main = paste(dat1$location[1],(dat1$year[1])),
     ylim = c(0,max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)), 
     xlim = c(0,max(PM_MB_means$AUDPS_m, na.rm = TRUE)))
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
points(grain_yield.t.ha. ~ AUDPS_m, data = dat1[dat1$fungicide_ai == "control",], col = "red")

@@

PM_MB_means$harvest_date - PM_MB_means$first_sign_disease
PM_MB_means[0 > as.Date(PM_MB_means$harvest_date, format = "%Y-%m-%d")-
as.Date(PM_MB_means$first_sign_disease, format = "%Y-%m-%d"),]

>>>>

The following code cnverts 'least square differences' (LSD) to sampling variance as per the method in [Nugugi et.al (2011)](https://apsjournals.apsnet.org/doi/10.1094/PHYTO-08-10-0221).
Ideas for using a T-critical value of 1.697 came from reading a [statisics-how-to website](https://www.statisticshowto.datasciencecentral.com/how-to-calculate-the-least-significant-difference-lsd/).
A [table of T-critical values](https://www.statisticshowto.datasciencecentral.com/tables/t-distribution-table/) was consulted where 30 degrees of freedom within experiment was used to find the 'T-crit' value of 1.697.

@@

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) # degrees of freedom within groups

   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error[1]

   V_yield <-
      (PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[1] * (LSD /
                                                                             Tcrit) ^ 2) / 2 # method - Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates))
      } # reset Sv vector if this is the first loop

@@

      Sv[j] = V_yield /
         PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j]

@@

   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$Y_error_type <-
      "SamplingVariance"

@@

for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error)) {
   if (j == 1) {
      Sv <-
         vector(length = length(PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error))
   }# create empty vector at first iteration with length of j

   Sv[j] = (PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error[j]) ^
      2
}

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$Y_error_type <-
   "SamplingVariance"
```

@@

for (i in unique(PM_MB_means$trial_ref)) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]

@@

   for (j in unique(dat1$row_spacing)) {
      dat2 <- dat1[dat1$row_spacing == j, ]
      controlY <-
         mean(dat2[dat2$fungicide_ai == "control", "grain_yield.t.ha."])

@@

      dat1[dat1$row_spacing == j, ] <- dat2
   }
   PM_MB_means[PM_MB_means$trial_ref == i, ] <- dat1
}

@@

dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303", ]

@@

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)
PM_MB_means[PM_MB_means$trial_ref == "AM1303", ] <- dat1

@@

```{r simpler_clusters_REMOVE}
slimmer_PM_dat <-
   slim_PM_dat[slim_PM_dat$spray_management != "Late" &
                  slim_PM_dat$spray_management != "Early_plus", ]

@@

slimmer_PM_dat %>%
   ggplot(aes(x = grain_yield.t.ha, y = vi)) +
   geom_point()
```

@@

```{r Var_MeanSquarePlot}
slimmer_PM_dat %>%
   ggplot(aes(x = grain_yield.t.ha, y = Y_Msquare)) +
   geom_point()
```

@@

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_means))

@@

```{r AUDPCxTrial}
sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPC_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPC_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)),
      xlim = c(0, max(PM_MB_means$AUDPC_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPC_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

@@

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_means))

@@

sm2lg <- PM_MB_means %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i,]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_means$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_means$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control", ],
          col = "red")

@@

PM_MB_means$harvest_date - PM_MB_means$first_sign_disease
PM_MB_means[0 > as.Date(PM_MB_means$harvest_date, format = "%Y-%m-%d")-
as.Date(PM_MB_means$first_sign_disease, format = "%Y-%m-%d"),]
```

### 2020-05-22 05_AUDPC_MetaAnalysis.Rmd

```{r AUDPC_funnel}
PM_Mdat$slope <- NA
PM_Mdat$SE <- NA
for(i in PM_Mdat$trial){
   PM_Mdat[PM_Mdat$trial == i,c("slope", "SE")] <-
      coef(summary(lm(grain_yield.t.ha ~ AUDPC_m,data = PM_Mdat[PM_Mdat$trial == i,])))[2,1:2]
}

plot(1/SE ~ slope, data = PM_Mdat, ylim = c(-3000,6000))
```

@@

```{r LMER_AUDPC1}
Me_mod1 <- lmer(grain_yield.t.ha ~ AUDPC_m + (1|trial),
              data = PM_Mdat)
summary(Me_mod1)
r.squaredGLMM(Me_mod1)

Me_mod1.s <- lmer(grain_yield.t.ha ~ AUDPS_m + (1|trial),
              data = PM_Mdat)
summary(Me_mod1.s)
r.squaredGLMM(Me_mod1)
r.squaredGLMM(Me_mod1.s)
anova(Me_mod1,Me_mod1.s)
```

@@

```{r LMER_AUDPC2}
Me_mod2 <- lmer(grain_yield.t.ha ~ AUDPC_m + (AUDPC_m|trial),
              data = PM_Mdat)
anova(Me_mod1, Me_mod2)

@@

```{r LMER_AUDPS2}
Me_mod2.s <- lmer(grain_yield.t.ha ~ AUDPS_m + (AUDPS_m|trial),
              data = PM_Mdat)
anova(Me_mod1.s, Me_mod2.s)

@@

Me_mod3 <- lmer(grain_yield.t.ha ~ PM_final_severity + (1|trial),
              data = PM_Mdat)
Me_mod4 <- lmer(grain_yield.t.ha ~ PM_final_severity + (PM_final_severity|trial),
              data = PM_Mdat)
anova(Me_mod3, Me_mod4)

@@

# PM_final_incidence vs AUDPC
anova(Me_mod4,Me_mod2)
```

@@

```{r AUDPC_spraymanagement}
Me_mod5 <- lmer(grain_yield.t.ha ~ AUDPC_m + spray_management+ (AUDPC_m|trial),
              data = PM_Mdat)

@@

```{r AUDPC:spraymangement}
Me_mod6 <- lmer(grain_yield.t.ha ~ AUDPC_m : spray_management+ (AUDPC_m|trial),
              data = PM_Mdat)

@@

```{r AUDPC:spraymangement2}
Me_mod8 <- lmer(grain_yield.t.ha ~ AUDPC_m : spray_management + (AUDPC_m: spray_management|trial),
              data = PM_Mdat)

@@

Me_mod9 <- lmer(grain_yield.t.ha ~ AUDPC_m : spray_management + (1|trial),
              data = PM_Mdat)

@@

```{r LMER_AUDPC2_results}
summary(Me_mod6)
coef(Me_mod6)[["trial"]][
   order(coef(Me_mod6)[["trial"]]$AUDPC_m),]

```
The first output shows our random effects do a good job of explaining the variation leaving less residual variation than what is explained by the random intercept, trial.
In the fixed effects the intercept gives an estimate of the average yield across all treatments (1.29 t/Ha). The other fixed effects show the change in predicted yield as AUDPC increases under each of the spray_management factors. Spray_management treatments, control, early, and recommended show significant declines in grain_yields as AUDPC increases. Recommended_plus and Late_plus show no significant decline in grain yield as AUDPC increases.

@@

PM_mv_AI <- rma.mv(yi,vi,
                   mods = ~ log(AUDPC_m+1):spray_management,
                   method = "ML",
                   random = list(~ spray_management | trial),
                   struct = "UN",
                   data = PM_Mdat)

summary(PM_mv_AI)

```
```{r}
exp(-0.1196)

>>>>

```{r AUDPC_funnel}
PM_Mdat$slope <- NA
PM_Mdat$SE <- NA
for (i in PM_Mdat$trial) {
   PM_Mdat[PM_Mdat$trial == i, c("slope", "SE")] <-
      coef(summary(lm(grain_yield.t.ha ~ AUDPC_m, data = PM_Mdat[PM_Mdat$trial == i, ])))[2, 1:2]
}

plot(1 / SE ~ slope,
     data = PM_Mdat,
     ylim = c(-3000, 6000))
```

@@

```{r LMER_AUDPC1}
Me_mod1 <- lmer(grain_yield.t.ha ~ AUDPC_m + (1 | trial),
                data = PM_Mdat)
summary(Me_mod1)
r.squaredGLMM(Me_mod1)

Me_mod1.s <- lmer(grain_yield.t.ha ~ AUDPS_m + (1 | trial),
                  data = PM_Mdat)
summary(Me_mod1.s)
r.squaredGLMM(Me_mod1)
r.squaredGLMM(Me_mod1.s)
anova(Me_mod1, Me_mod1.s)
```

@@

```{r LMER_AUDPC2}
Me_mod2 <- lmer(grain_yield.t.ha ~ AUDPC_m + (AUDPC_m | trial),
                data = PM_Mdat)
anova(Me_mod1, Me_mod2)

@@

```{r LMER_AUDPS2}
Me_mod2.s <- lmer(grain_yield.t.ha ~ AUDPS_m + (AUDPS_m | trial),
                  data = PM_Mdat)
anova(Me_mod1.s, Me_mod2.s)

@@

Me_mod3 <- lmer(grain_yield.t.ha ~ PM_final_severity + (1 | trial),
                data = PM_Mdat)
Me_mod4 <-
   lmer(grain_yield.t.ha ~ PM_final_severity + (PM_final_severity |
                                                   trial),
        data = PM_Mdat)
anova(Me_mod3, Me_mod4)

@@

# PM_final_incidence vs AUDPC
anova(Me_mod4, Me_mod2)
```

@@

```{r AUDPC_spraymanagement}
Me_mod5 <-
   lmer(grain_yield.t.ha ~ AUDPC_m + spray_management + (AUDPC_m |
                                                            trial),
        data = PM_Mdat)

@@

```{r AUDPC:spraymangement}
Me_mod6 <-
   lmer(grain_yield.t.ha ~ AUDPC_m:spray_management + (AUDPC_m | trial),
        data = PM_Mdat)

@@

```{r AUDPC:spraymangement2}
Me_mod8 <-
   lmer(
      grain_yield.t.ha ~ AUDPC_m:spray_management + (AUDPC_m:spray_management |
                                                        trial),
      data = PM_Mdat
   )

@@

Me_mod9 <-
   lmer(grain_yield.t.ha ~ AUDPC_m:spray_management + (1 | trial),
        data = PM_Mdat)

@@

```{r LMER_AUDPC2_results}
summary(Me_mod6)
coef(Me_mod6)[["trial"]][order(coef(Me_mod6)[["trial"]]$AUDPC_m),]
```
The first output shows our random effects do a good job of explaining the variation leaving less residual variation than what is explained by the random intercept, trial.
In the fixed effects the intercept gives an estimate of the average yield across all treatments (1.29 t/Ha). The other fixed effects show the change in predicted yield as AUDPC increases under each of the spray_management factors. Spray_management treatments, control, early, and recommended show significant declines in grain_yields as AUDPC increases. Recommended_plus and Late_plus show no significant decline in grain yield as AUDPC increases.

@@

PM_mv_AI <- rma.mv(
   yi,
   vi,
   mods = ~ log(AUDPC_m + 1):spray_management,
   method = "ML",
   random = list( ~ spray_management | trial),
   struct = "UN",
   data = PM_Mdat
)

summary(PM_mv_AI)
```

```{r}
exp(-0.1196)
